
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(GGally)
library(stargazer)
library(here)
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(kableExtra)
library(broom)
library(ggbeeswarm)
library(effsize)

```

# Overview of Mono Lake Levels from 1850 to 2017 and Recent History
### By Morganne Sigismonti
```{r}
# Read in the data and cleaned up
mono_lake_clean <- read_csv(here("data", "mono_lake_annual_levels.csv"), skip = 5) %>%
  clean_names() %>% 
  rename("lake_level" = lake_level_feet_above_sea_level) %>%
  select(year, lake_level)
  
 
```

```{r}
# Created a plot of the 
mono_lake_clean %>%
  ggplot(aes(x = year, y = lake_level, fill = lake_level)) +
  geom_line(size = .8, color = "dodgerblue2", alpha = .5) +
  scale_x_continuous(limits = c(1850, 2018.3), # made sure I could see 2017 in the graph
                     breaks = c(1850, 1875, 1900, 1925, 1950, 1975, 2000, 2017), #years I want the graph to display
                     expand = c(0, 0)) +
  geom_ribbon(aes(ymin = -Inf, ymax = lake_level),
              fill = "dodgerblue2",
              alpha = .3) + #added fill under the line 
  theme_minimal() +
  labs(title = "Change in Mono Lake Levels, 1850 - 2017 (SIGISMONTI)\n",x = "\nYear", y = "Lake Level (ft above sea level)\n") +
  
  # added annotations to the historically significant years.
  #1941
  annotate(geom="text", x=1941, y= 6422,
           label="1941", size = 2.75) +
  annotate(geom= "point", x=1941, y= 6417, color = "black") +
  
  #1979
  annotate(geom="text", x=1979, y= 6369,
           label="1979", size = 2.75) +
  annotate(geom= "point", x=1979, y= 6373.40, color = "black") +
  
  #1984 diversion stopped Mono lake streams and lake 
  annotate(geom="text", x=1984, y= 6386,
           label="1984", size = 2.75) +
  annotate(geom= "point", x=1984, y= 6380.10, color = "black") +
  
  #1994
  annotate(geom="text", x=1994, y= 6370,
           label="1994", size = 2.75) +
  annotate(geom= "point", x=1994, y= 6374.60, color = "black") +
  
  #1998
  annotate(geom="text", x=1998, y= 6390,
           label="1998", size = 2.75) +
  annotate(geom= "point", x=1998, y= 6384.30, color = "black") +
  
  #2013
  annotate(geom="text", x=2013, y= 6387,
           label="2013", size = 2.75) +
  annotate(geom= "point", x=2013, y= 6380.60, color = "black") +
  geom_hline(yintercept=6377, color = "green4") +
  geom_hline(yintercept=6360, color = "red") +
  geom_hline(yintercept=6350, color = "black")


```

**Figure 1** Water levels (ft above sea level) for Mono Lake from 1850 through 2017 with important historical points and critical water levels. Black points, indicate years that influenced water levels in Mono Lake. In 1941, the Los Angeles Department of Water & Power (DWP) began diverting water from Mono Lake’s tributary streams. In 1979, Mono Lake Committee (MLC) sued the DWP. In 1984, the MLC had won multiple cases against the DWC and water began flowing down the Mono Basin streams’ dry channels. In 1994, decision 1631 set a lake level was made to keep Mono Lake at 6,392 feet above sea level and ordered DWP to restore streams and waterfowl habitat damaged from diversions. In 1998, two restoration orders, 98-05 and 98-07, were issued by the State Water Board issued, requiring DWP to implement a restoration plan for Mono Basin's streams and waterfowl habitat. In 2013, responsible parties joined the Mono Basin Stream Restoration Agreement. The green line (6,377 ft above sea level) indicates when land bridges emerge and allow predators to reach major bird nesting sites, the red line (6,360 ft above sea level) indicates when water would have a salinity of 120 g/L and cause a decrease in brine shrimp availability for Eared Grebes and California Gulls, and the black line (lake level of 6,360 ft) indicates when water would have a salinity above 150 g/L and cause acute impacts.
