---
title: "a6_task2_morganne_sigismonti"
author: "Morganne Sigismonti"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(GGally)
library(stargazer)
library(here)
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(kableExtra)
```


```{r, include=FALSE}
# Read the data in, cleaned it up, and made a table with political party choice and environmentalist identity choice
conservation_env_pol <- read_csv(here("data", "conservation_survey.csv")) %>%
  clean_names() %>% 
  select(env_id, pol_or) %>% #only include these columns from the data
  mutate(cons_lib = case_when( #add a new column that bins political party
    pol_or %in% c(1,2,3) ~ "Conservative", 
    pol_or %in% c(5,6,7) ~ "Liberal")) %>%
      filter(!is.na(cons_lib)) %>% #exclude other data
  mutate(environmentalist = case_when( #add a new column that bins responses about if a person is an environmentalist
    env_id == "1" ~ "Yes",
     env_id == "2" ~ "No",
     env_id == "3" ~ "Unsure"))

```

```{r, include=FALSE}
# Made a table comparing political party and environmentalist identity choice including counts and percentages
survey_results <- conservation_env_pol %>%
  tabyl(cons_lib,environmentalist) %>%
  adorn_percentages() %>%
  janitor::adorn_pct_formatting(digits = 2) %>%
  adorn_ns() %>% 
 relocate("cons_lib", "Yes", "No", "Unsure")
  

```


**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for weights of female and male juvenile snowshoe hares. Data: Kielland et al. (2017)
```{r}
survey_results %>%
   kable(col.names = c("", "Yes", "No", "Unsure")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
```
## Make contingency table and run chi-square

First, we see that the site is its own variable - when we run chi-square, we really only want the counts in the table. 

We can use `column_to_rownames` to convert entries in a column to stored rownames: 

```{r}
# Made a contingency table with just the counts of environmentalist idenitity choice with political party choice
survey_cont_table <- conservation_env_pol %>%
  tabyl(cons_lib,environmentalist) %>% 
  column_to_rownames(var = "cons_lib") %>%
  relocate("Yes", "No", "Unsure")


```


```{r}
#performed a chi-square test for independence to answer: is there an association between political affiliation (conservative or liberal-leaning) and if a person considers theirself an environmentalist? 

survey_results_x2 <- chisq.test(survey_cont_table)
survey_results_x2

survey_results_tidy <- tidy(survey_results_x2)
```

