---
title: "a6_task3_morganne_sigismonti"
author: "Morganne Sigismonti"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(GGally)
library(stargazer)
library(here)
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(kableExtra)
library(broom)
library(ggbeeswarm)
```

For Task 3, you are asked to compare total lengths (variable = total_length, in millimeters) for female lizards (sex == “F”) of two species (variable spp): CNUN (Cnemidophorus uniparens, the Desert Grassland lizard) and CNTE (Cnemidophorus tessalatus, the Colorado Checkered lizard).


```{r, include = FALSE}
# Create and store a simplified version of the data frame that only includes variables spp, sex, and total_length
# Then, filtered to only include observations for female lizards of species CNUN and CNTE

lizards_data <- read_csv(here("data", "lizards.csv")) %>%
  clean_names() %>% 
  select(spp, sex, total_length) %>% 
  filter(spp %in% c("CNUN","CNTE")) %>% 
  filter(sex == "F") %>% 
  filter(!is.na(sex)) %>% 
  filter(!is.na(total_length))
```

```{r}
# make a beeswarm plot with box plot, mean, and median of female lizard lengths for each species (CNUN and CNTE)


ggplot(data = lizards_data, aes(x = spp, y = total_length)) +
  geom_beeswarm(aes(color = spp, fill = spp)) +
  geom_boxplot(fill = NA, width = 0.2, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point",
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  theme_classic() +
  labs(title = "Desert Grassland and Colorado Checkered Female Lizard Lengths (SIGISMONTI)\n",
       x = "\nSpecies",
       y = "Total Length (mm)") +
   scale_color_discrete()


```
### Figure Caption
**Figure 1.**

Prepare a finalized summary table that contains the mean and standard deviation for total length, and sample size, for each group (female CNUN and CNTE lizards)

# WHAT ABOUT NA VALUES???


**Table 1.** Mean, standard deviation, and sample size for each species, CNTE (Colorado Checkered lizard) and CNUN (Desert Grassland lizard).
```{r}
## make a table 
lizards_stats <- lizards_data %>%
 group_by("Species" = spp) %>% 
  summarize("Mean" = mean(total_length),
            "Standard Deviation" = sd(total_length),
            "Sample Size" = n())


kbl(lizards_stats)%>%
 kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
```{r}
# Exploratory Analysis

## histogram/qq plot
## decide on parametric/nonparametric test to run - see notes to decide which to choose 
## get cohens d
### write a thing about it
```


## Analysis

Is there a significant difference in total length difference in mean total length between female lizards for the two species?
