
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(GGally)
library(stargazer)
library(here)
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(kableExtra)
library(broom)
```

# An Anaysis of Association of Political Affiliation with Environmental Identity
### By Morganne Sigismonti

## Introduction

The data used in this study was collected by Chelsea Batavia and Andrews Forest LTER from 1,331 anonymous, nationwide U.S. respondents through an online survey conducted in August, 2017. (Batavia et al. (2019)) In this analysis, I will compare responses for political affiliation (Conservative or Liberal) and self-identification as an environmentalist (Yes, No, Unsure) to attempt to find out if these two variables are associated with one-another. I will set up a contingency table and use the Chi-Squared test for independence to compare expected counts (if there was no affect of political affiliation on self-identification as an environmentalist ) and actual counts to see if these variables are independent or affiliated.


```{r, include=FALSE}
# Read the data in, cleaned it up, and made a table with political party choice and environmental identity choice
conservation_env_pol <- read_csv(here("data", "conservation_survey.csv")) %>%
  clean_names() %>% 
  select(env_id, pol_or) %>% #only include these columns from the data
  mutate(cons_lib = case_when( #add a new column that bins political party
    pol_or %in% c(1,2,3) ~ "Conservative", 
    pol_or %in% c(5,6,7) ~ "Liberal")) %>%
      filter(!is.na(cons_lib)) %>% #exclude other data
  mutate(environmentalist = case_when( #add a new column that bins responses about if a person is an environmentalist
    env_id == "1" ~ "Yes",
     env_id == "2" ~ "No",
     env_id == "3" ~ "Unsure"))

```

```{r, include=FALSE}
# Made a table comparing political party and environmental identity choice including counts and percentages
survey_results <- conservation_env_pol %>%
  tabyl(cons_lib,environmentalist) %>%
  adorn_percentages() %>%
  janitor::adorn_pct_formatting(digits = 2) %>%
  adorn_ns() %>% 
 relocate("cons_lib", "Yes", "No", "Unsure")
  

```


**Table 1.** Counts and proportions of Political Affiliation (Conservative or Liberal) and Environmentalist Identity (Yes, No, or Unsure) Data: Batavia et al. (2019)
```{r}
survey_results %>%
   kable(col.names = c("", "Yes", "No", "Unsure")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
```

```{r}
# Made a contingency table with just the counts of environmentalist idenitity choice with political party choice
survey_cont_table <- conservation_env_pol %>%
  tabyl(cons_lib,environmentalist) %>% 
  column_to_rownames(var = "cons_lib") %>%
  relocate("Yes", "No", "Unsure")


```

```{r, include = FALSE}
# Performed a chi-square test for independence to answer: is there an association between political affiliation (conservative or liberal-leaning) and if a person considers themselves an environmentalist? 

survey_results_x2 <- chisq.test(survey_cont_table)
survey_results_x2

survey_results_tidy <- tidy(survey_results_x2)

```
## Results and Takeaways
There is a significant association between political affiliation and if a person considers themselves an environmentalist ($\chi$^2^(`r survey_results_tidy$parameter`) = `r round(survey_results_tidy$statistic,2)`, *p* = `r format(survey_results_tidy$p.value, scientific = TRUE, digits = 3)`. We have significant evidence to reject the null hypothesis of independence. This mean that we retain the alternative hypothesis that political affiliation and environmentalist identity are not independent. This can be seen in the contingency table, where 53.73% (281 people) of Liberal identifying respondents identified as environmentalists and only 33.65% (213) of Conservative identifying respondents identified as environmentalists.This shows that this data overall shows 


## Citation
Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. 

